Seg_sites_rarefaction

### 2024-04-09 scripts/seg_sites_rarefaction.Rmd
supplementary data and snpR per-locus for SI

# Rarefact via simulations
Note: this chunck should run on a cluster or somewhere else allowing for large numbers of threads, then be re-imported in the next chunck. `q` should be passed as an argument for the iteration, as should `outfile`.

```{r cluster_MCMCs, eval=FALSE}
set.seed(seed + q) # reset seed for cluster

@@

# draw for each k/mcmc count combination
combs <- expand.grid(k, mcmcs)
colnames(combs) <- c("k", "iters")

tmcmcs <- combs$iters[q]
tk <- combs$k[q]

seg1 <- seg2 <- matrix(0, nal, tmcmcs)
for(i in 1:tmcmcs){
  for(j in 1:nal){
    seg1[j,i] <- rarefact_mcmc(g1[j,], k = tk)
    seg2[j,i] <- rarefact_mcmc(g2[j,], k = tk)
  }
}

mpseg1 <- binom::binom.confint(rowSums(seg1), rep(tmcmcs, nrow(seg1)),
                               method = "agresti-coull")[,c("mean", "lower", "upper")]
mpseg2 <- binom::binom.confint(rowSums(seg2), rep(tmcmcs, nrow(seg1)),
                               method = "agresti-coull")[,c("mean", "lower", "upper")]

colnames(mpseg1) <- colnames(mpseg2) <- c("mean", "lower", "upper")

res <- list(seg1 = seg1, seg2 = seg2, mpseg1 = mpseg1, mpseg2 = mpseg2, k = tk, iters = tmcmcs)

saveRDS(res, paste0(outfile, "_", q, ".RDS"))

@@

>>>>

# Rarefact via simulations
Note: this chunk should run on a cluster or somewhere else allowing for large numbers of threads, then be re-imported in the next chunk. `q` should be passed as an argument for the iteration, as should `outfile`.

```{r cluster_MCMCs, eval=FALSE}
# set.seed(seed + q) # reset seed for cluster

@@

# # draw for each k/mcmc count combination
# combs <- expand.grid(k, mcmcs)
# colnames(combs) <- c("k", "iters")
# 
# tmcmcs <- combs$iters[q]
# tk <- combs$k[q]
# 
# seg1 <- seg2 <- matrix(0, nal, tmcmcs)
# for(i in 1:tmcmcs){
#   for(j in 1:nal){
#     seg1[j,i] <- rarefact_mcmc(g1[j,], k = tk)
#     seg2[j,i] <- rarefact_mcmc(g2[j,], k = tk)
#   }
# }
# 
# mpseg1 <- binom::binom.confint(rowSums(seg1), rep(tmcmcs, nrow(seg1)),
#                                method = "agresti-coull")[,c("mean", "lower", "upper")]
# mpseg2 <- binom::binom.confint(rowSums(seg2), rep(tmcmcs, nrow(seg1)),
#                                method = "agresti-coull")[,c("mean", "lower", "upper")]
# 
# colnames(mpseg1) <- colnames(mpseg2) <- c("mean", "lower", "upper")
# 
# res <- list(seg1 = seg1, seg2 = seg2, mpseg1 = mpseg1, mpseg2 = mpseg2, k = tk, iters = tmcmcs)
# 
# saveRDS(res, paste0(outfile, "_", q, ".RDS"))

@@

#========real data example, mathematically============

```{r}
library(snpR)
set.seed(12321)
d <- readRDS("../data/monarchs_non_strict.RDS")
d <- d[sample(nrow(d), 1000),] # subsample 1k loci

# remove smaller than 10, although it works fine including stuff like SAI with only 4
sample.meta(d)$pop[sample.meta(d)$pop %in% c("ENA", "WNA")] <- "NAM"
ps <- summarize_facets(d, "pop")$pop
d <- d[pop = -c(names(ps)[which(ps < 10)])]

d <- calc_seg_sites(d, "pop", g = 0) # set gamma to the smallest N per loci
seg_res_mon <- get.snpR.stats(d, "pop", "seg_sites") # follows biogeographic history
```

#=========real data example, via rarefaction==========
```{r}
# setup
sn <- format_snps(d, "sn", interpolate = FALSE)
sn <- sn[,-c(1:7)]
tmcmcs <- 100

pops <- substr(colnames(sn), 1, 3)
upops <- unique(pops)
mon_seg <- array(0, c(nrow(sn), tmcmcs, length(unique(pops))))

# do the rarefaction
for(i in 1:tmcmcs){
  cat("\nIter:", i)
  for(j in 1:nrow(sn)){
    # cat("\n\t", j)
    tg <- min(tapply(unlist(sn[j,]), pops, function(x) sum(!is.na(x)))) # set gamma for this locus
    for(k in 1:length(upops)){
      mon_seg[j,i,k] <- rarefact_mcmc(sn[j,which(pops == upops[k])], k = tg) # rarefact this locus in this pop
    }
  }
}

# summarize results
mon_rare_res <- matrix(0, tmcmcs, length(upops))
for(k in 1:length(upops)){
  mon_rare_res[,k] <- colSums(mon_seg[,,k])
}
colnames(mon_rare_res) <- upops
mon_rare_res <- as.data.table(mon_rare_res)
mon_rare_res$i <- 1:tmcmcs
mon_rare_res <- melt(mon_rare_res, id.vars = "i", variable.name = "pop", value = "Ns")

ggplot(mon_rare_res, aes(x = pop, y = Ns)) +
  geom_boxplot() +
  geom_point(alpha = .2, position = position_jitter(height = 0, width = .2)) +
  geom_point(data = seg_res_mon$weighted.means, aes(x = subfacet, y = seg_sites), color = "red") +
  theme_bw() +
  ylab(bquote(italic(N[s]))) +
  xlab("Population")
